The hidden manifold distance for functional data

### 2019-06-26 index.Rmd
fixed reg_sampling to mean concentrated sampling
more paper writing

---
title: "The hidden manifold distance for functional manifold data"
output:
  html_document:
    df_print: paged
    number_sections: true
---

@@

1. Transform each vector $Y_i$ into a function $\tilde X_i$ by spline smoothing:
$$ \tilde X_i = \arg\min_{f\in C^2[0,1]}\left\{\sum_{j=1}^{K}\left(f(t_{ij})-Y_{ij}\right)^2+\lambda \|\partial^2_tf\|^2_{L^2}\right\}$$
where $\lambda>0$ is a tuning parameter controlling the smoothness of $\tilde X_i$. 
2. Obtain a $s$-dimensional representation $\tilde X^s_1,\ldots,\tilde X^s_n$ of the functions $\tilde X_1,\ldots,\tilde X_n$ that "preserves" the pairwise $L^2$ distances by using MDS.
3. Obtain $\tilde X^{s,\hat \M}_1,\ldots,\tilde X^{s,\hat \M}_n$ a "projection" of $\tilde X^s_1,\ldots,\tilde X^s_n$ onto a ridge $\hat \M$ which is computed with a mean shift algorithm.
4. Use  IsoGeo to approximate the pairwise geodesic distances $\{d_{\hat \M}(\tilde X^{s,\hat \M}_i,\tilde X^{s,\hat \M}_j)\}_{i>j}$ and define the $n \times n $ matrix $\hat G$ as
$$
\hat G(i,j)=\hat G(j,i) = \left\{ \begin{array}{ll}
 d_{\hat \M}(\tilde X^{s,\hat \M}_i,\tilde X^{s,\hat \M}_j) & \textrm{if $i\neq j$,}\\
 0 & \textrm{otherwise.}
  \end{array} \right.
$$

@@

## Simulation scenarios

@@

## Results

# Study on "misspecification"

@@

# Downstream tasks

In this section, we explore whether our geodesic distance estimator has benefits for downstream analysis task. There are many tasks we could consider here, but we will focus on distanced-based functional clustering and distance-based functional classification. It must be noted that while curve alignment, also known as curve registration, is necessarily performed as a preprocessing technique prior to clustering and classification, our geodesic distance estimator allows one to forsake this step.

In both tasks below, we compare our method to using $L_2$ distance, possibly weighted, and with curve registration already accomplished.

## Distance-based functional clustering

Describe the $k$-medoids clustering method.

### Simulated data

###  Real data

## Distance-based functional classification

For simplicity, assume the task is bianry classification. Associated to each functional object $x$ is a binary $y$ indicating class membership.
Consider the classifier proposed in Ferraty and Vieu (2003,2006) which is a functional version of the Nadaraya-Watson kernel estimator of class membership probabilities:
$$
\hat p(y = 0 | x) \frac{ \sum_{i=1}^n K[h^{-1} d(x,x_i)] 1(y_i = 0) }{ \sum_{i=1}^n K[h^{-1} d(x,x_i)] }
$$

### Simulated data

###  Real data
Datasets used by functional classification papers

@@

>>>>

---
title: "The hidden manifold distance for functional data"
output:
  html_document:
    df_print: paged
    number_sections: yes
  pdf_document: default
---

@@

1. Transform each vector $Y_i$ into a function $\tilde X_i$ by spline smoothing:
$$ \tilde X_i = \arg\min_{f\in C^2[0,1]}\left\{\sum_{j=1}^{K}\left(f(t_{ij})-Y_{ij}\right)^2+\lambda \|\partial^2_tf\|^2_{L^2}\right\}$$
where $\lambda>0$ is a tuning parameter controlling the smoothness of $\tilde X_i$. 
2. Obtain a $s$-dimensional representation $\tilde X^s_1,\ldots,\tilde X^s_n$ of the functions $\tilde X_1,\ldots,\tilde X_n$ that "preserves" the pairwise $L^2$ distances by using MDS.
3. Obtain $\tilde X^{s,\hat \M}_1,\ldots,\tilde X^{s,\hat \M}_n$ a "projection" of $\tilde X^s_1,\ldots,\tilde X^s_n$ onto a ridge $\hat \M$ which is computed with a mean shift algorithm.
4. Use  IsoGeo to approximate the pairwise geodesic distances $\{d_{\hat \M}(\tilde X^{s,\hat \M}_i,\tilde X^{s,\hat \M}_j)\}_{i>j}$ and define the $n$ by $n$ matrix 
$$
\hat G(i,j)=\hat G(j,i) = \left\{ \begin{array}{ll}
 d_{\hat \M}(\tilde X^{s,\hat \M}_i,\tilde X^{s,\hat \M}_j) & \textrm{if $i\neq j$,}\\
 0 & \textrm{otherwise.}
  \end{array} \right.
$$

@@

@@

## Simulation scenarios
Acknowledge that we are sampling on a very concentrated measure to make sure functional manifold is well sampled. Cite Diaconis et. al's "Sampling from a manifold". How to sample properly from a functional manifold could be interesting future work.

* warping manifold (scenario 5 in sim_functional_data.R)
* pdf manifold (scenario 2 in sim_functional_data.R)
* square root pdf manifold (scenario 4 )
### Results

Submit compare_methods.sh as slurm job using sbatch.
```{r, engine = 'bash', eval = FALSE}
sbatch compare_methods.sh
```

Alternatively, we can run the local version
```{r, eval=FALSE}
R compare_methods_local.R
```

Results are visualized by invoking cluster_results.R
```{r, eval=FALSE}
R cluster_results.R
```

## The cost of employing proposed method when manifold is flat

@@

# Distance-based functional classification

In this section, we explore whether our geodesic distance estimator has benefits for downstream analysis task. There are many tasks we could consider here such as distance-based nonparametric regression and distanced-based functional clustering, but we will focus on distance-based functional classification. It must be noted that while curve alignment, also known as curve registration, is necessarily performed as a preprocessing technique prior to clustering and classification, our geodesic distance estimator allows one to forsake this step.

For simplicity, assume the task is bianry classification. Associated to each functional object $x$ is a binary $y$ indicating class membership.
Consider the classifier proposed in Ferraty and Vieu (2003,2006) which is a functional version of the Nadaraya-Watson kernel estimator of class membership probabilities:
$$
\hat p(y = 0 | x) \frac{ \sum_{i=1}^n K[h^{-1} d(x,x_i)] 1(y_i = 0) }{ \sum_{i=1}^n K[h^{-1} d(x,x_i)] }
$$
We shall compare our method to using $L_2$ distance, possibly weighted, and with curve registration already accomplished.

Datasets used by functional classification papers

@@

<!-- ## Distance-based functional clustering -->

<!-- Describe the $k$-medoids clustering method. -->

<!-- ### Simulated data -->

<!-- ###  Real data -->
